\documentclass[a4paper]{article}

\usepackage{amsmath}
\usepackage{geometry}

\begin{document}

\begin{itemize}
\item Susceptible individuals [$S$] are exposed to the virus by
  contact with an infected individual.
\item Exposed individuals [$E$] progress towards a symptomatic state
  on average time $t_l$
\item Infected individuals [$I$] infect an average of $R_0$ secondary
  infections. On a time-scale of $t_i$, infected individuals either
  recover or progress towards hospitalization.
\item Hospitalized individuals [$H$] either recover or worsen towards
  a critical state on a time-scale of $t_h$.
\item Critical individuals [$C$] model ICU usage. They either return
  to the hospital state or die [$D$] on a time-scale of $t_c$
\item Overflow individuals [$O$] model those that need ICU treatment
  but do not receive it. They either return to the hospital state or
  die [$D$] on a time-scale of $t_o$
\item Recovered individuals [$R$] can not be infected again.
\end{itemize}

\begin{align*} 
  \frac{dS_{a}(t)}{dt} & = -N^{-1}\beta_a(t) S_a(t)\sum_{b}I_b(t) \\ 
  \frac{dE_{a}(t)}{dt} & = N^{-1}\beta_a(t) S_a(t)\sum_{b}I_b(t) - E_{a}(t)/t_l \\ 
  \frac{dI_{a}(t)}{dt} & = E_{a}(t)/t_l - I_a(t)/t_i \\ 
  \frac{dH_{a}(t)}{dt} & = (1-m_a) I_a(t)/t_i + (1-f_a) C_a(t)/t_c + (1 - g_a) O_a(t)/t_o - H_a(t)/t_h \\ 
  \frac{dC_{a}(t)}{dt} & = c1_a H_a(t)/t_h - C_a(t)/t_c \\ 
  \frac{dO_{a}(t)}{dt} & = c0_a H_a(t)/t_h - O_a(t)/t_o \\ 
  \frac{dR_{a}(t)}{dt} & = m_a I_a(t)/t_i + (1-c_a)H_a(t)/t_h \\ 
  \frac{dD_{a}(t)}{dt} & = f_a C_a(t)/t_c + g_a O_a(t)/t_o
\end{align*}

Overflow survival rate $g_a = \gamma f_a/(\gamma f_a + 1 - f_a)$
where $\gamma$ is the death odds multiplier when critical and not
receiving ICU treatment.

The critical to ICU rate $c1_a$ and the critical to overflow rate $c0_a$ are:
\begin{align*}
	c0_a &= c_a\sigma\left (\sum_{b=0}^a C_b(t) - \lambda\right ) \\
	c1_a &= c_a - c0_a
\end{align*}
where $\lambda$ is the number of ICU beds and $\sigma()$ is the sigmoid function.
The sum of the critical count up to $a$ ensures that younger people take preference when deciding who gets the ICU beds.
The sigma acts as a switch; $c0_a$ is zero, until the critical count passes $\lambda$, after which $c1_a$ becomes zero.

The transmission rate $\beta_a(t)$ is given by
\begin{align*}
\beta_a(t) = R_0 \zeta_a M(t) (1+\varepsilon \cos(2\pi (t-t_{max})))/t_i
\end{align*}


\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:
